All figures
========================================================

```{r load, message=FALSE, warning=FALSE, echo=FALSE}
# Here's how to compile this file in R, assuming your current directory is the repository root:

# library(knitr)
# knit("./reports/AllFigures.Rmd", encoding="utf-8")

library(affy)
library(ggplot2)
library(reshape)
library(plyr)
library(cluster)
library(xtable)
library(MASS)

options(xtable.type="html")
```

```{r loadexpr, echo=FALSE}
load("../cache/eset.original.filter.RData")
dataset <- eset.original.filter
```


```{r dopca, message=FALSE, warning=FALSE, echo=FALSE}
pca <- prcomp(t(exprs(dataset)), scale=TRUE, center=TRUE)
pc <- pca$x
```

PCA plot on MCC/SCLC tumor samples and MCC cell lines, using PC1 and PC2
--------------------

```{r pcascatter, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=10}
## Plot on PC1 and PC2
pc.data <- data.frame(pc, sample=rownames(pc))
rownames(pc.data) <- rownames(pc)
pc.data <- cbind(pc.data[rownames(pc), ], pData(dataset)[rownames(pc), ])

p <- ggplot(pc.data, aes(x=PC1, y=PC2)) + geom_point(aes(color=cancertype, 
                                                         shape=factor(sample.type, 
                                                                      labels=c("Cell line", "Tissue"))),
                                                     size=6)
p <- p + xlim(-75, 200)
p <- p + geom_text(aes(label=rownames(pc.data)), hjust=0.8, vjust=-0.75, size=4)
p <- p + scale_colour_brewer("Cancer type", palette="Set1")
p <- p + scale_shape_manual("Sample type", values=c(18,16))
p <- p + labs(x="Principal component 1", y="Principal component 2")
p <- p + theme_bw() + theme(legend.position=c(0.85, 0.85),
                           legend.title=element_text(size=14),
                           legend.text=element_text(size=12),
                           legend.background=element_rect(colour="black"),
                           axis.title.x=element_text(size=14), axis.title.y=element_text(size=14, angle=90),
                           axis.text.x=element_text(size=12), axis.text.y=element_text(size=12),
                           panel.background=element_rect())

print(p)

ggsave("../graphs/pc1v2.pdf", p, width=10, height=10)

```

Variance plot of PCA on MCC and SCLC tumor samples and cell lines
----------------------

```{r pcavariance, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10}
## Proportion of variance
pca.summary <- summary(pca)
pca.var <- data.frame(variance=pca.summary$importance[2, ], pc=names(pca.summary$importance[2, ]))
pca.var <- head(pca.var, 10)
pca.var <- transform(pca.var, pc=factor(pc, levels=names(pca.summary$importance[2, ]), ordered=TRUE))
p <- ggplot(pca.var, aes(x=pc, y=variance)) + geom_bar(stat="identity") + labs(x="Principal Component", y="Variance")
p <- p + theme_bw() + theme(axis.text.x=element_text(size=10, angle=270), 
                            axis.text.y=element_text(size=10),
                            legend.position="none") ## + coord_flip()

print(p)
ggsave("../graphs/pc1v2.var.pdf", p, width=8, height=8)

```


Correlation box plot of MCC/SCLC tumor samples and MCC cell lines
-------------------

```{r corrboxplot, message=FALSE, warning=FALSE, echo=FALSE, fig.width=15, fig.height=7}

dataset.original <- eset.original.filter

# Subset of samples and make sample classes
which.MCCtumor <- grep("MT.*", sampleNames(dataset.original))
which.mkl1 <- grep("Mkl.*", sampleNames(dataset.original))
which.waga <- grep("Waga.*", sampleNames(dataset.original))
which.uiso <- grep("UISO.*", sampleNames(dataset.original))

keep <- c(which.MCCtumor, which.mkl1, which.waga, which.uiso)
dataset.original <- dataset.original[, keep]

dataset <- dataset.original

# Compute correlation
dataset.cor <- cor(exprs(dataset), method='spearman')

# Keep only the necessary columns and rows and
# set self correlation values (including within group) to NA
dataset.cor[upper.tri(dataset.cor, diag=TRUE)] <- NA
dataset.cor[grep("Mkl", rownames(dataset.cor)), grep("Mkl", rownames(dataset.cor))] <- NA
dataset.cor[grep("Waga", rownames(dataset.cor)), grep("Waga", rownames(dataset.cor))] <- NA
dataset.cor[grep("UISO", rownames(dataset.cor)), grep("UISO", rownames(dataset.cor))] <- NA

exclude.rows <- -(1:(length(which.MCCtumor)))
dataset.cor <- dataset.cor[exclude.rows, ]

# Reshape the data
dataset.cor <- transform(dataset.cor, sample=rownames(dataset.cor))
dataset.cor.melted <- melt(dataset.cor, id.vars=c("sample"), na.rm=TRUE)

# For some reason, melt changed the Mkl-1's name...
dataset.cor.melted$variable <- factor(gsub("\\.", "-", dataset.cor.melted$variable))

# Get sample_class from the phenotypeData based on the sample column
dataset.cor.melted <- merge(dataset.cor.melted, pData(dataset)[, c("sample", "class")], by.x="sample", by.y="sample")

dataset.cor.melted$class <- factor(dataset.cor.melted$class, 
                                   levels=c("Mkl1", "WaGa", "UISO"),
                                   ordered=TRUE)

# Get everything from the phenotypeData again, based on the variable column
dataset.cor.melted <- merge(dataset.cor.melted, pData(dataset.original), by.x="variable", by.y="sample")
dataset.cor.melted <- dataset.cor.melted[, c('sample', 'variable', 'value', 'class.x', "class.y", "MCPyV.status")]

# Add some grouping factors
dataset.cor.melted <- transform(dataset.cor.melted,
                                comparison=factor(with(dataset.cor.melted, paste(class.x, class.y, sep=".")),
                                                  levels=c("Mkl1.Tumor", "WaGa.Tumor", "UISO.Tumor", "WaGa.Mkl1", "UISO.Mkl1", "UISO.WaGa"),
                                                  ordered=TRUE),
                                tumorORnot=factor(dataset.cor.melted$class.y == "Tumor", #c("MCC Tumor", "SCLC Tumor"), 
                                                  labels=c("Cell line vs. cell line", "Cell line vs. tumor")))


# plot
p <- ggplot(dataset.cor.melted, aes(x=comparison, y=value))
p <- p + geom_boxplot(notch=FALSE, outlier.size=0)
p <- p + geom_jitter(aes(color=MCPyV.status, shape=MCPyV.status), position=position_jitter(width=0.1), alpha=0.6)
p <- p + facet_grid(. ~ tumorORnot, scales="free_x")
p <- p + labs(x="", y=expression(Correlation~(rho)))
p <- p + scale_color_brewer(palette='Set1')
p <- p + guides(colour=guide_legend("Tumor MCV Status"), shape=guide_legend("Tumor MCV Status"))
p <- p + ylim(c(0.5,1))
p <- p + theme_bw() + theme(axis.text=element_text(size=20), 
                            legend.title=element_text(size=20),
                            legend.text=element_text(size=20),
                            legend.position=c(0.6, 0.2),
                            axis.title.x=element_text(size=20),
                            axis.title.x=element_text(size=20),
                            axis.title.y=element_text(size=20, angle=90),
                            strip.text.x=element_text(size=24))

print(p)

ggsave("../graphs/corrboxplot.pdf", p, width=16, height=8)

```                        

Hierarchical clustering on MCC/SCLC tumor samples and MCC cell lines
------------------------

```{r hierclust, message=FALSE, warning=FALSE, echo=FALSE, fig.width=12, fig.height=6}
dataset <- eset.original.filter
d <- as.dist(1 - cor(exprs(dataset), method="spearman"))
plot(agnes(d, diss=TRUE, method='average'), which.plots=2, xlab="Dissimilarity = 1 - Correlation", main="Average Linkage")

pdf("../graphs/hier.pdf", width=12, height=6)
plot(agnes(d, diss=TRUE, method='average'), which.plots=2, xlab="Dissimilarity = 1 - Correlation", main="Average Linkage")
dev.off()

```

Random forest using MCC/SCLC tumor data applied to MCC/UISO lines
========================================================

```{r loadrf, echo=FALSE, error=FALSE, warning=FALSE, cache=FALSE, message=FALSE}
load("../cache/rf.eset.RData")
load("../cache/rf.predict.RData")
load("../cache/rf.eset.unsupervised.RData")
load("../cache/classes.train.RData")
load("../cache/classes.train.unsupervised.RData")
load("../cache/samplenames.train.RData")
load("../cache/samplenames.train.unsupervised.RData")
```

## Cancer type error rate of random forest on MCC/SCLC tumor samples (using out-of-bag samples)
```{r echo=FALSE, results='asis'}
tmp <- data.frame(table(classes.train))
class.err.rate <- as.numeric(tail(rf.eset$err.rate,1))[-1]
tbl <- data.frame(Frequency=tmp$Freq, Error=class.err.rate)
rownames(tbl) <- tmp$classes.train
print(xtable(tbl))
```

## Average test set predictions of random forest applied to MCC cell lines
``` {r testerror, echo=FALSE, results='asis'}
rf.predict.mean <- ddply(rf.predict, "class", function(x) data.frame(MCC=mean(x$MCC), SCLC=mean(x$SCLC)))
rf.predict.mean.xtable <- xtable(rf.predict.mean)
print(rf.predict.mean.xtable)
```

## Multi-Dimensional scaling plot of (unsupervised) random forest proximities on MCC/SCLC tumor samples and MCC cell lines
```{r mdsrfallcelllines, fig.width=12, fig.height=12, echo=FALSE}

rf.mds <- cmdscale(1 - rf.eset.unsupervised$proximity, eig=TRUE)

plotdata <- data.frame(x=rf.mds$points[,1], y=rf.mds$points[,2],
                       labels=samplenames.train.unsupervised, cancertype=classes.train.unsupervised)
p <- ggplot()

p <- p + geom_point(data=plotdata, aes(x=x, y=y, 
                                       color=cancertype),
                    size=5, show_guide=TRUE)

p <- p + scale_colour_brewer("Cancer type", palette="Set1")
p <- p + scale_shape_manual("Sample type", values=c(18,16))

p <- p + geom_text(data=plotdata, aes(x=x, y=y, label=rownames(plotdata)),
                   hjust=1.0, vjust=-1.25, size=4)

p <- p + labs(x="Scaling component 1", y="Scaling component 2")
p <- p + theme_bw() + theme(legend.position=c(0.85,0.5),
                            legend.title=element_text(size=14),
                            legend.text=element_text(size=12),
                            legend.background=element_rect(colour="black"),
                            panel.background = element_rect(fill = "#FFFFFF"))
print(p)

# for publication
ggsave("figure/mdsmccsclc.pdf", p, width=8, height=8)
```

Commands to run GSEA using GSEAPreranked on the output from limma's moderated t-test
------------------------

Unfortunately, this step is difficult to automatically reproduce because GSEA writes out files to a folder with a timestamp appended. All files run with these parameters are provided.

The following code will reproduce these files, but the timestamp of the output folder will be different, and you would need to change that in the following code manually.

From the command line, assumes that your current directory is the repository root:

```
sh ./data/GSEA/run_GSEA.sh
```

```{r readdata, echo=FALSE}
data <- list()
data[['tumor.classic.pos']] <- read.table("../data/GSEA/tumor_vs_classic_filter/gsea_report_for_na_pos.xls", sep="\t", header=TRUE)
data[['tumor.classic.pos']]$direction <- 1

data[['tumor.classic.neg']] <- read.table("../data/GSEA/tumor_vs_classic_filter/gsea_report_for_na_neg.xls", sep="\t", header=TRUE)
data[['tumor.classic.neg']]$direction <- -1

data[['classic.variant.pos']] <- read.table("../data/GSEA/classic_vs_variant_filter/gsea_report_for_na_pos.xls", sep="\t", header=TRUE)
data[['classic.variant.pos']]$direction <- 1

data[['classic.variant.neg']] <- read.table("../data/GSEA/classic_vs_variant_filter/gsea_report_for_na_neg.xls", sep="\t", header=TRUE)
data[['classic.variant.neg']]$direction <- -1

data[['tumor.variant.pos']] <- read.table("../data/GSEA/tumor_vs_variant_filter/gsea_report_for_na_pos.xls", sep="\t", header=TRUE)
data[['tumor.variant.pos']]$direction <- 1

data[['tumor.variant.neg']] <- read.table("../data/GSEA/tumor_vs_variant_filter/gsea_report_for_na_neg.xls", sep="\t", header=TRUE)
data[['tumor.variant.neg']]$direction <- -1

merged.data <- ldply(data, function(x) x[, c("NAME", "FDR.q.val", "direction")])

## Process data
merged.data$NAME <- gsub("KEGG_", "", merged.data$NAME)
merged.data$.id <- gsub("\\.pos|\\.neg", "", merged.data$.id)
merged.data$FDR.q.val <- round(merged.data$FDR.q.val, 6)

# fdr.thresh.init <- 0.25
# merged.data <- subset(merged.data, FDR.q.val <= fdr.thresh.init)
# merged.data <- subset(merged.data, FDR.q.val <= 0.95 & FDR.q.val >= 0.9)

merged.data$FDR.q.val.formatted <- format.pval(merged.data$FDR.q.val, digits=4)

merged.data$value <- ifelse(merged.data$direction == 1,
                            gsub("(.*)", '<div style="color: red;">\\1</div>', merged.data$FDR.q.val),
                            gsub("(.*)", '<div style="color: green;">\\1</div>', merged.data$FDR.q.val))

fdr.thresh <- 0.05
merged.data$value <- ifelse(merged.data$FDR.q.val <= fdr.thresh,
                            gsub("(.*)", '<div style="font-size:150%">\\1</div>', merged.data$value),
                            merged.data$value)

  
merged.data.wide <- cast(merged.data, NAME ~ .id, value="value")
rownames(merged.data.wide) <- merged.data.wide$NAME
merged.data.wide <- merged.data.wide[, -1]
merged.data.wide <- as.data.frame(merged.data.wide)

merged.data.wide.fdr <- cast(merged.data, NAME ~ .id, value="FDR.q.val")
rownames(merged.data.wide.fdr) <- rownames(merged.data.wide)
merged.data.wide.fdr <- merged.data.wide.fdr[, -1]

merged.data.wide.fdr <- as.data.frame(apply(merged.data.wide.fdr, 2, as.numeric))
rownames(merged.data.wide.fdr) <- rownames(merged.data.wide)
colnames(merged.data.wide.fdr) <- colnames(merged.data.wide)

## Anything that is NA is FDR = 1
merged.data.wide.fdr[is.na(merged.data.wide.fdr)] <- 1

merged.data.wide <- merged.data.wide[order(merged.data.wide.fdr$tumor.variant), ]

merged.data.wide$classic.variant <- as.character(merged.data.wide$classic.variant)
merged.data.wide$tumor.variant <- as.character(merged.data.wide$tumor.variant)
merged.data.wide$tumor.classic <- as.character(merged.data.wide$tumor.classic)

set.na <- is.na(merged.data.wide.fdr)
merged.data.wide[set.na] <- "NS" # paste("> ", fdr.thresh, sep="")

merged.data.wide <- merged.data.wide[, c("tumor.variant", "tumor.classic", "classic.variant")]

colnames(merged.data.wide) <- c( "UISO vs. Tumor", "Classic vs. Tumor", "UISO vs. Classic")

```

## Results for GSEA, differentially expressed in the cell lines (number of pathways: `r nrow(merged.data.wide)`):

To identify pathways that are affected similarly, we performed three comparisons: UISO cell line vs. tumor samples, classic cell lines vs. tumor samples, and UISO cell line vs. classic cell lines. Green values are up-regulated and red values are down-regulated. Large bold values are statistically significant (FDR-adjusted p-value < `r fdr.thresh`).

```{r table.all, echo=FALSE, results='asis'} 
print(xtable(merged.data.wide, digits=c(1,2,2,2), align=c("r", "c", "c", "c")), type='html', sanitize.text.function=I) 
``` 


```{r loadccle, echo=FALSE, error=FALSE, warning=FALSE, cache=FALSE, message=FALSE}
load("../cache/rf.eset.allcelllines.RData")
load("../cache/rf.eset.nouiso.predict.RData")
load("../cache/rf.eset.nouiso.RData")

classes.train.allcelllines <- rf.eset.allcelllines$y
samplenames.train.allcelllines <- names(rf.eset.allcelllines$predicted)

```

## Prediction on UISO cell line samples using RF trained with CCLE + MCC cell lines
```{r predictuiso, echo=FALSE, results='asis'}
print(xtable(rf.eset.nouiso.predict, digits=2))
```

## Cancer Type Error Rate with CCLE + MCC cell lines
```{r errorratesnouiso, echo=FALSE, results='asis'}
tmp <- data.frame(table(rf.eset.nouiso$y))
class.err.rate <- as.numeric(tail(rf.eset.nouiso$err.rate,1))[-1]
tbl <- data.frame(Frequency=tmp$Freq, Error=class.err.rate)
rownames(tbl) <- as.character(levels(rf.eset.nouiso$y))
print(xtable(tbl))
# print(xtable(tbl))
```

## Confusion Matrix with CCLE + MCC cell lines
```{r confusionmatrixnouiso, echo=FALSE, results='asis'}
print(xtable(rf.eset.nouiso$confusion, digits=c(rep(0, 10), 2)))

``` 

## Average Out-Of-Bag Votes (Per Class) with CCLE + MCC cell lines
```{r avgvotestblnouiso, echo=FALSE, results='asis'}
oobvotes <- data.frame(rf.eset.nouiso$votes, cancertype=rf.eset.nouiso$y)
avgoobvotes <- ddply(oobvotes, .variables="cancertype", function(x) colMeans(x[, -ncol(x)]))
rownames(avgoobvotes) <- avgoobvotes[, 1]
avgoobvotes <- avgoobvotes[, -1]
colnames(avgoobvotes) <- levels(rf.eset.nouiso$y)
print(xtable(avgoobvotes, digits=2))
```

## Cancer Type Error Rate with all cell lines (MCC, UISO, CCLE)
```{r errorratesccle, echo=FALSE, results='asis'}
tmp <- data.frame(table(rf.eset.allcelllines$y))
class.err.rate <- as.numeric(tail(rf.eset.allcelllines$err.rate,1))[-1]
tbl <- data.frame(Frequency=tmp$Freq, Error=class.err.rate)
rownames(tbl) <- as.character(levels(rf.eset.allcelllines$y))
print(xtable(tbl))
```

## Confusion Matrix with all cell lines (MCC, UISO, CCLE)
```{r confusionmatrixccle, echo=FALSE, results='asis'}
print(xtable(rf.eset.allcelllines$confusion, digits=c(rep(0, 10), 2)))
``` 

## Average Out-Of-Bag Votes (Per Class) with all cell lines (MCC, UISO, CCLE)
```{r avgvotestblccle, echo=FALSE, results='asis'}
oobvotes <- data.frame(rf.eset.allcelllines$votes, cancertype=rf.eset.allcelllines$y)
avgoobvotes <- ddply(oobvotes, .variables="cancertype", function(x) colMeans(x[, -ncol(x)]))
rownames(avgoobvotes) <- avgoobvotes[, 1]
avgoobvotes <- avgoobvotes[, -1]
colnames(avgoobvotes) <- levels(rf.eset.allcelllines$y)
print(xtable(avgoobvotes, digits=2))
```


## Multi-Dimensional Scaling Plot of all cell lines (MCC, UISO, CCLE)
```{r mdsrfallcelllinesccle, fig.width=12, fig.height=12, echo=FALSE, message=FALSE, warning=FALSE}
rf.mds <- cmdscale(1 - rf.eset.allcelllines$proximity, add=TRUE, eig=TRUE)

plotdata <- data.frame(x=rf.mds$points[,1], y=rf.mds$points[,2],
                       labels=samplenames.train.allcelllines, cancertype=classes.train.allcelllines)

mcc.uisos <- grepl("UISO", rownames(plotdata))

plotdata$cancertype <- as.character(plotdata$cancertype)
plotdata$cancertype[mcc.uisos] <- "UISO"
plotdata$cancertype <- factor(plotdata$cancertype)

mcc.uisos <- grep("UISO", plotdata$labels, value=FALSE)

plotdata <- transform(plotdata, predicted=rf.eset.allcelllines$predicted)
plotdata <- transform(plotdata, correct=factor((as.character(predicted) == as.character(cancertype)), levels=c(TRUE, FALSE), ordered=TRUE))

plotdata.other <- subset(plotdata, !(cancertype %in% c('MCC', 'UISO')))
plotdata.MCC <- subset(plotdata, cancertype == 'MCC')
plotdata.UISO <- subset(plotdata, cancertype == 'UISO')

legend.labels <- paste(levels(classes.train.allcelllines), " (", table(classes.train.allcelllines), ")", sep="")
legend.labels <- legend.labels[-c(6,10)]

p <- ggplot()
p <- p + geom_point(data=plotdata.other, aes(x=x, y=y, color=cancertype), alpha=0.5, size=5, show_guide=TRUE)

p <- p + scale_colour_manual("Cancer type",
                            values=c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#D94801", "#A65628", "#F781BF"),
                            labels=legend.labels) 

p <- p + geom_point(data=plotdata.MCC, aes(x=x, y=y), size=7, color="black")
p <- p + geom_point(data=plotdata.MCC, aes(x=x, y=y), size=5, color="#F41A1C")
p <- p + geom_text(data=plotdata.MCC, aes(x=x, y=y, label=rownames(plotdata.MCC)),
                   hjust=1.0, vjust=-1.25, size=4)

p <- p + geom_point(data=plotdata.UISO, aes(x=x, y=y), size=7, color="black")
p <- p + geom_point(data=plotdata.UISO, aes(x=x, y=y), size=5, color="#6BAED6")
p <- p + geom_text(data=plotdata.UISO, aes(x=x, y=y, label=rownames(plotdata.UISO)),
                   hjust=-0.1, vjust=-0.75, size=4)

p <- p + labs(x="Scaling component 1", y="Scaling component 2")
p <- p + theme_bw() + theme(legend.position=c(0.1, 0.85),
                            legend.title=element_text(size=14),
                            legend.text=element_text(size=12),
                            legend.background=element_rect(colour="black"),
                            panel.background = element_rect(fill = "#FFFFFF"))
print(p)

# for publication
ggsave("figure/mdsrfallcelllines.pdf", p, width=8, height=8)
```

PCA plot of CCLE, MCC, and UISO cel lines using PC1 and PC2
--------------------
``````{r pcaccle, message=FALSE, warning=FALSE, echo=FALSE, cache=TRUE, fig.width=12, fig.height=12}
load("../cache/eset.batcheffect.nomod.RData")

dataset <- eset.batcheffect.nomod
pca <- prcomp(t(exprs(dataset)), scale=TRUE, center=TRUE)
pc <- pca$x

plotdata <- data.frame(pc, sample=rownames(pc))
rownames(plotdata) <- rownames(pc)
plotdata <- cbind(plotdata[rownames(pc), ], pData(dataset)[rownames(pc), ])

```

```{r pcascatterccle, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=10}
## Plot on PC1 and PC2

plotdata$plot.type <- plotdata$cancer.type

mcc.uisos <- grepl("UISO", rownames(plotdata))

plotdata$plot.type <- as.character(plotdata$plot.type)
plotdata$plot.type[mcc.uisos] <- "UISO"
plotdata$plot.type <- factor(plotdata$plot.type)

mcc.uisos <- grep("UISO", plotdata$labels, value=FALSE)

plotdata.other <- subset(plotdata, !(plot.type %in% c('MCC', 'UISO')))
plotdata.MCC <- subset(plotdata, plot.type == 'MCC')
plotdata.UISO <- subset(plotdata, plot.type == 'UISO')

legend.labels <- levels(plotdata$plot.type)
legend.labels <- legend.labels[-c(6,10)]

p <- ggplot()
p <- p + geom_point(data=plotdata.other, aes(x=PC1, y=PC2, color=plot.type), alpha=0.5, size=5, show_guide=TRUE)

p <- p + scale_colour_manual("Cancer type",
                            values=c("#E41A1C", "#377EB8", "#4DAF4A", "#984EA3", "#FF7F00", "#D94801", "#A65628", "#F781BF"),
                            labels=legend.labels) 

p <- p + geom_point(data=plotdata.MCC, aes(x=PC1, y=PC2), size=7, color="black")
p <- p + geom_point(data=plotdata.MCC, aes(x=PC1, y=PC2), size=5, color="#F41A1C")
p <- p + geom_text(data=plotdata.MCC, aes(x=PC1, y=PC2, label=rownames(plotdata.MCC)),
                   hjust=1.0, vjust=-1.25, size=4)

p <- p + geom_point(data=plotdata.UISO, aes(x=PC1, y=PC2), size=7, color="black")
p <- p + geom_point(data=plotdata.UISO, aes(x=PC1, y=PC2), size=5, color="#6BAED6")
p <- p + geom_text(data=plotdata.UISO, aes(x=PC1, y=PC2, label=rownames(plotdata.UISO)),
                   hjust=-0.1, vjust=-0.75, size=4)

p <- p + labs(x="Principal component 1", y="Principal component 2")
p <- p + theme_bw() + theme(legend.position=c(0.85, 0.85),
                            legend.title=element_text(size=14),
                            legend.text=element_text(size=12),
                            legend.background=element_rect(colour="black"),
                            panel.background = element_rect(fill = "#FFFFFF"))
print(p)

ggsave("../graphs/pc1v2ccle.pdf", p, width=12, height=12)

```
