Supplementary Tables and Figures
-------------------

## 

```{r load, message=FALSE, warning=FALSE, echo=FALSE}
# Here's how to compile this file in R, assuming your current directory is the
# repository root:

# library(rmarkdown)
# render("./reports/SuppMat.Rmd", html_document(css="style.css"))

# Or, like this:

# library(knitr)
# knit("./reports/SuppMat.Rmd", encoding="utf-8")
 
library(affy)
library(ggplot2)
library(reshape)
library(plyr)
library(cluster)
library(xtable)
library(MASS)
library(limma)
library(knitr)
library(RColorBrewer)

source("./lib.R")
source("../lib/gsea_helpers.R")

options(xtable.type="html", xtable.caption.placement="top", xtable.include.row.names=FALSE)
opts_chunk$set(comment=NA, echo=FALSE, warning=FALSE, message=FALSE)

```

```{r loadexpr, echo=FALSE}
load("../cache/eset.original.filter.RData")
dataset <- eset.original.filter
```

### Figure S0: Scatterplot matrix of first three principal components.

```{r figs0, fig.width=10, fig.height=10}
pca <- prcomp(t(exprs(dataset)), scale=TRUE, center=TRUE)
pc <- pca$x

plotdata <- data.frame(pc, sample=rownames(pc))
rownames(plotdata) <- rownames(pc)
plotdata <- cbind(plotdata[rownames(pc), ], pData(dataset)[rownames(pc), ])

plotdata$newclass <- ifelse(plotdata$class == "Tumor",
                           as.character(plotdata$cancertype),
                           as.character(plotdata$class))

plotdata$newclass <- factor(plotdata$newclass, ordered=TRUE,
                           levels=c("MC01", "WaGa", "Mkl1", "UISO", "MCC13", "MCC26", "MCC", "SCLC"),
                           labels=c("MC01", "WaGa", "Mkl-1", "UISO", "MCC13", "MCC26", "MCC Tumor", "SCLC Tumor"))

# expand iris data frame for pairs plot
gg1 = makePairs(plotdata[, c("PC1", "PC2", "PC3")])

# new data frame mega iris
mega_data <- data.frame(gg1$all,
                       class=rep(plotdata$newclass, length=nrow(gg1$all)),
                       sample.type=rep(plotdata$sample.type, length=nrow(gg1$all)),
                       MCPyV.status=rep(plotdata$MCPyV.status, length=nrow(gg1$all)))

# pairs plot
type.colors <- c(brewer.pal(9, "Blues")[c(8, 4, 6)], 
                 brewer.pal(9, "Greens")[c(8)], 
                 brewer.pal(9, "Oranges")[c(4, 6)], 
                 brewer.pal(9, "Blues")[c(9)], 
                 brewer.pal(9, "Reds")[c(8)]) 

point.shapes <- c(19,19,19,19,19,19,17,17)

p <- ggplot(mega_data, aes_string(x = "x", y = "y"))
p <- p + facet_grid(xvar ~ yvar, scales = "free")
p <- p + geom_point(aes(colour=class, shape=sample.type), na.rm = TRUE, alpha=0.8, size=4)

# MCV positive cases
mega_data.MCV <- subset(mega_data, MCPyV.status == "Virus negative")
p <- p + geom_point(data=mega_data.MCV, colour="white", shape="x", na.rm = TRUE, alpha=0.8, size=4)

# p <- p + stat_density(aes(x = x, y = ..scaled.. * diff(range(x)) + min(x)),
#                       data = gg1$densities, position = "identity",
#                       colour = "grey20", geom = "line")
p <- p + scale_color_manual("Sample group", values=type.colors)
p <- p + scale_shape_manual("Sample type", values=c(19,17), guide = 'none')
p <- p + guides(color=guide_legend("Sample group", override.aes = list(shape=point.shapes)))
p <- p + labs(x="", y="")
p <- p + theme_bw() + theme(legend.position="top",
                            strip.text=element_text(size=14),
                            axis.title=element_text(size=14),
                            axis.text=element_blank(),
                            legend.title=element_text(size=14),
                            legend.text=element_text(size=12))
p

```

<div style="page-break-after:always"></div>

### Figure S1: The classic cell lines are more similar to MCC tumor samples than the UISO cell line.

Boxplots of pairwise Spearman correlations of global, variance-filtered probeset expression signatures between different cell line samples (WaGa, Mkl-1 and UISO) and between cell line and tumor samples. For each box: middle bar, median; box, inter-quartile range (middle 50% of data); bars extend to 1.5 times the inter-quartile range. The color of individual points indicates the MCV status of the tumor samples.

```{r prepfigS1, message=FALSE, warning=FALSE, echo=FALSE}

# MCC samples only, don't use MC01 since it's a singleton

keep <- with(pData(eset.original.filter), which(cancertype == "MCC" & sample != "MC01"))
dataset.original <- eset.original.filter[, keep]
dataset <- dataset.original

# Compute correlation
dataset.cor <- cor(exprs(dataset), method='spearman')

# Keep only the necessary columns and rows and
# set self correlation values (including within group) to NA
dataset.cor[upper.tri(dataset.cor, diag=TRUE)] <- NA

dataset.cor[grep("MCC", rownames(dataset.cor)),
            grep("MCC", rownames(dataset.cor))] <- NA
dataset.cor[grep("Mkl", rownames(dataset.cor)),
            grep("Mkl", rownames(dataset.cor))] <- NA
dataset.cor[grep("Waga", rownames(dataset.cor)),
            grep("Waga", rownames(dataset.cor))] <- NA
dataset.cor[grep("UISO", rownames(dataset.cor)),
            grep("UISO", rownames(dataset.cor))] <- NA

which.MCCtumor <- grep("MT", rownames(dataset.cor))

exclude.rows <- -(1:(length(which.MCCtumor)))
dataset.cor <- dataset.cor[exclude.rows, ]

# Reshape the data
dataset.cor <- transform(dataset.cor, sample=rownames(dataset.cor))
dataset.cor.melted <- melt(dataset.cor, id.vars=c("sample"), na.rm=TRUE)

# For some reason, melt changed the Mkl-1's name...
dataset.cor.melted$variable <- factor(gsub("\\.", "-", 
                                           dataset.cor.melted$variable))

# Get sample_class from the phenotypeData based on the sample column
dataset.cor.melted <- merge(dataset.cor.melted,
                            pData(dataset)[, c("sample", "classic")],
                            by.x="sample", by.y="sample")

dataset.cor.melted$classic <- factor(dataset.cor.melted$classic, 
                                   levels=c("ClassicCellLine", "VariantCellLine", "UISOCellLine"),
                                   # labels=c("WaGa", "Mkl-1", "UISO"),
                                   ordered=TRUE)

# Get everything from the phenotypeData again, based on the variable column
dataset.cor.melted <- merge(dataset.cor.melted,
                            pData(dataset),
                            by.x="variable", by.y="sample")

dataset.cor.melted <- dataset.cor.melted[, c('sample', 'variable', 'value', 'classic.x', "classic.y", "MCPyV.status")]

# Add some grouping factors
dataset.cor.melted <- transform(dataset.cor.melted,
                                comparison=factor(paste(classic.x, classic.y, sep=".")),
                                tumorORnot=factor(dataset.cor.melted$classic.y == "Tumor",
                                                  labels=c("Cell line vs. cell line", "Cell line vs. tumor")))

dataset.cor.melted$comparison <- gsub("CellLine", "", dataset.cor.melted$comparison)

dataset.cor.melted$finalcomp <- NULL

classicvsvariant <- with(dataset.cor.melted, 
                         comparison %in% c("Classic.Variant", "Variant.Classic"))
dataset.cor.melted$finalcomp[classicvsvariant] <- "ClassicVariant"

classicvsuiso <- with(dataset.cor.melted, 
                      comparison %in% c("Classic.UISO", "UISO.Classic"))
dataset.cor.melted$finalcomp[classicvsuiso] <- "ClassicUISO"

variantvsuiso <- with(dataset.cor.melted, 
                      comparison %in% c("Variant.UISO", "UISO.Variant"))
dataset.cor.melted$finalcomp[variantvsuiso] <- "VariantUISO"

classicvsclassic <- with(dataset.cor.melted, 
                      comparison %in% c("Classic.Classic"))
dataset.cor.melted$finalcomp[classicvsclassic] <- "ClassicClassic"

dataset.cor.melted$finalcomp <- ifelse(is.na(dataset.cor.melted$finalcomp),
                                       as.character(dataset.cor.melted$comparison),
                                       dataset.cor.melted$finalcomp)

```

```{r figS1, message=FALSE, warning=FALSE, echo=FALSE, fig.width=10, fig.height=8}

# plot
p <- ggplot(dataset.cor.melted, aes(x=finalcomp, y=value))
p <- p + geom_boxplot(notch=FALSE, outlier.size=0)
p <- p + geom_jitter(aes(color=MCPyV.status, shape=MCPyV.status), position=position_jitter(width=0.1), alpha=0.8)
p <- p + facet_grid(. ~ tumorORnot, scales="free_x")
p <- p + labs(x="", y=expression(Correlation~(rho)))
p <- p + scale_color_brewer(palette='Set1')
p <- p + guides(colour=guide_legend("Tumor MCV Status"), shape=guide_legend("Tumor MCV Status"))
p <- p + ylim(c(0.5,1))
p <- p + theme_bw() + theme(axis.text.x=element_text(size=18, angle=270, vjust=0.5), 
                            axis.text.y=element_text(size=18), 
                            legend.title=element_text(size=20),
                            legend.text=element_text(size=18),
                            legend.position="top",
                            axis.title.x=element_text(size=20),
                            axis.title.y=element_text(size=20, angle=90),
                            strip.text.x=element_text(size=20))

print(p)

ggsave("../graphs/figS1.pdf", p, width=16, height=8)

```

<div style="page-break-after:always"></div>

### Table S1: Differential expression of MCC tumor samples using clinical parameters.

The number of probesets that passed significance thresholds for pairwise differential expression tests (fold change > 2 and q-value < 0.05) and the minimum q-value observed in each pairwise differential expression test.

```{r TableS1, results='asis', echo=FALSE}
load("../cache/fit2.tumors.MCPyV.status.RData")
load("../cache/fit2.tumors.privsmet.RData")
load("../cache/fit2.tumors.recurrence.RData")
load("../cache/fit2.tumors.stage.code.RData")

fit2.list <- list(MCPyV.status=fit2.tumors.MCPyV.status,
                  privsmet=fit2.tumors.privsmet,
                  recurrence=fit2.tumors.recurrence,
                  stage.code=fit2.tumors.stage.code)

fit2.nrows <- ldply(fit2.list, function(x) nrow(topTable(x, number=Inf, lfc=1, p=0.05)))
fit2.minqval <- ldply(fit2.list, function(x) min(topTable(x, number=Inf)$adj.P.Val))

class1 <- c("Positive (20)", "Primary (10)", "No (7)", "Early (5)")
class2 <- c("Negative (3)", "Metastasis (13)", "Yes (16)", "Late (18)")
fit2.data <- data.frame(class1=class1, class2=class2, number=fit2.nrows$V1, minqval=fit2.minqval$V1)

rownames(fit2.data) <- c("MCV Status", "Primary vs. Metastasis", "Recurrence", "Stage")

colnames(fit2.data) <- c("Class 1", "Class 2", "# Significant Probesets", "Minimum q-value")
fit2.xtbl <- xtable(fit2.data, align=c('r', 'r', 'r', 'c', 'c'))
print(fit2.xtbl, sanitize.text.function=I, include.rownames=TRUE)
write.csv(fit2.data, file="tables/tables1.csv")

```

### Table S2: Median correlation values of cell line-tumor comparisons between Merkel cell polyomavirus positive and negative tumor samples.

```{r TableS2, echo=FALSE, results='asis', include=TRUE}
# Show the median correlation values by mcv status
mcv.corr <- ddply(dataset.cor.melted, .(comparison, MCPyV.status), function(x) data.frame(median=median(x$value)))
mcv.corr <- mcv.corr[grep("Tumor", mcv.corr$comparison), ]
mcv.corr$comparison <- factor(mcv.corr$comparison)
mcv.corr <- transform(mcv.corr, comparison=factor(comparison, labels=gsub("\\.", " vs. ", levels(comparison))))
mcv.corr.cast <- cast(mcv.corr, comparison ~ MCPyV.status, value="median")
rownames(mcv.corr.cast) <- mcv.corr.cast$comparison
mcv.corr.cast <- mcv.corr.cast[1:3, -1]
mcv.corr.xtbl <- xtable(mcv.corr.cast)
print(mcv.corr.xtbl, include.rownames=TRUE)
write.csv(mcv.corr.cast, file="tables/tables2.csv")
```

<div style="page-break-after:always"></div>

### Table S3: Gene Set Enrichment Analysis (GSEA) on KEGG pathways using differentially expressed genes in the cell lines

```{r readdata, echo=FALSE}
# See the following script to reproduce all provided files (assumes that your current directory is the repository root):
# sh ./data/GSEA/run_GSEA.sh

# Samples annotated as pos or 1 are those that are 
data <- list()

tumor.classic.template <- "../data/GSEA/tumor_vs_classic_filter.GseaPreranked"
tumor.classic.run <- "1408460601222"
tumor.classic.pos.file <- gsea_result_filename(tumor.classic.template, tumor.classic.run, "pos")
tumor.classic.neg.file <- gsea_result_filename(tumor.classic.template, tumor.classic.run, "neg")

data[['tumor.classic.pos']] <- read.table(tumor.classic.pos.file, sep="\t", header=TRUE)
data[['tumor.classic.pos']]$direction <- 1

data[['tumor.classic.neg']] <- read.table(tumor.classic.neg.file, sep="\t", header=TRUE)
data[['tumor.classic.neg']]$direction <- -1

tumor.uiso.template <- "../data/GSEA/tumor_vs_uiso_filter.GseaPreranked"
tumor.uiso.run <- "1408460539704"
tumor.uiso.pos.file <- gsea_result_filename(tumor.uiso.template, tumor.uiso.run, "pos")
tumor.uiso.neg.file <- gsea_result_filename(tumor.uiso.template, tumor.uiso.run, "neg")

data[['tumor.uiso.pos']] <- read.table(tumor.uiso.pos.file, sep="\t", header=TRUE)
data[['tumor.uiso.pos']]$direction <- 1

data[['tumor.uiso.neg']] <- read.table(tumor.uiso.neg.file, sep="\t", header=TRUE)
data[['tumor.uiso.neg']]$direction <- -1

tumor.variant.template <- "../data/GSEA/tumor_vs_variant_filter.GseaPreranked"
tumor.variant.run <- "1408460663055"
tumor.variant.pos.file <- gsea_result_filename(tumor.variant.template, tumor.variant.run, "pos")
tumor.variant.neg.file <- gsea_result_filename(tumor.variant.template, tumor.variant.run, "neg")

data[['tumor.variant.pos']] <- read.table(tumor.variant.pos.file, sep="\t", header=TRUE)
data[['tumor.variant.pos']]$direction <- 1

data[['tumor.variant.neg']] <- read.table(tumor.variant.neg.file, sep="\t", header=TRUE)
data[['tumor.variant.neg']]$direction <- -1

merged.data <- ldply(data, function(x) x[, c("NAME", "FDR.q.val", "direction")])

## Process data
merged.data$NAME <- gsub("KEGG_", "", merged.data$NAME)
merged.data$.id <- gsub("\\.pos|\\.neg", "", merged.data$.id)
merged.data$FDR.q.val <- round(merged.data$FDR.q.val, 6)


merged.data$FDR.q.val.formatted <- format.pval(merged.data$FDR.q.val, digits=1)

merged.data$value <- ifelse(merged.data$direction == 1,
                            gsub("(.*)",
                                 # '<div style="color: red;">\\1</div>',
                                 '<div class="red"><span>\\1</span></div>',
                                 merged.data$FDR.q.val.formatted),
                            gsub("(.*)", 
                                 # '<div style="color: green;">\\1</div>',
                                 '<div class="green"><span>\\1</span></div>',
                                 merged.data$FDR.q.val.formatted))

fdr.thresh <- 0.05
merged.data$value <- ifelse(merged.data$FDR.q.val <= fdr.thresh,
                            gsub("(.*)",
                                 '<div style="font-size:150%">\\1</div>',
                                 merged.data$value),
                            merged.data$value)

  
merged.data.wide <- cast(merged.data, NAME ~ .id, value="value")
rownames(merged.data.wide) <- merged.data.wide$NAME
merged.data.wide <- merged.data.wide[, -1]
merged.data.wide <- as.data.frame(merged.data.wide)

merged.data.wide.fdr <- cast(merged.data, NAME ~ .id, value="FDR.q.val")
rownames(merged.data.wide.fdr) <- rownames(merged.data.wide)
merged.data.wide.fdr <- merged.data.wide.fdr[, -1]

merged.data.wide.fdr <- as.data.frame(apply(merged.data.wide.fdr, 2, 
                                            as.numeric))
rownames(merged.data.wide.fdr) <- rownames(merged.data.wide)
colnames(merged.data.wide.fdr) <- colnames(merged.data.wide)

## Anything that is NA is FDR = 1
merged.data.wide.fdr[is.na(merged.data.wide.fdr)] <- 1

merged.data.wide <- merged.data.wide[order(merged.data.wide.fdr$tumor.uiso), ]

merged.data.wide$tumor.uiso <- as.character(merged.data.wide$tumor.uiso)
merged.data.wide$tumor.variant <- as.character(merged.data.wide$tumor.variant)
merged.data.wide$tumor.classic <- as.character(merged.data.wide$tumor.classic)

set.na <- is.na(merged.data.wide.fdr)
merged.data.wide[set.na] <- "NS"

merged.data.wide <- merged.data.wide[, c("tumor.uiso", "tumor.variant", "tumor.classic")]
merged.data.wide.fdr <- merged.data.wide.fdr[order(merged.data.wide.fdr$tumor.uiso), ]
merged.data.wide.fdr <- merged.data.wide.fdr[, c("tumor.uiso", "tumor.variant", "tumor.classic")]

colnames(merged.data.wide) <- c( "UISO vs. Tumor", "Variant vs. Tumor",
                                 "Classic vs. Tumor")

```

To identify KEGG pathways (n = `r nrow(merged.data.wide)`) that are affected similarly, we performed three comparisons: UISO cell line vs. tumor samples, classic cell lines vs. tumor samples, and UISO cell line vs. classic cell lines. Green values are up-regulated and red values are down-regulated. Large bold values are statistically significant (FDR-adjusted p-value < `r fdr.thresh`).


```{r TableS3, echo=FALSE, results='asis'}
tmp <- xtable(merged.data.wide, digits=c(1,2,2,2), align=c("r", "c", "c", "c"))
print(tmp, type='html', sanitize.text.function=I)
write.csv(merged.data.wide, file="tables/tables3.csv")

``` 


<div style="page-break-after:always"></div>

### Table S4: Variant MCC cell line samples are not classified as MCC with a random forest classifier trained using cell line microarray expression data.

The mean prediction votes of variant (MCC13, MCC26, and UISO) and classic (Mkl-1 and WaGa; labeled MCC) cell line samples are given.
^* The MCC samples are used for training so bias may exist.
See Table S5 for unbiased prediction estimates of Mkl-1 and WaGa samples.

```{r loadccle, echo=FALSE, error=FALSE, warning=FALSE, cache=FALSE, message=FALSE}
load("../cache/rf.eset.nouiso.predict.RData")
load("../cache/rf.eset.nouiso.predict.train.RData")
load("../cache/rf.eset.nouiso.RData")
```

```{r TableS4, echo=FALSE, results='asis'}
rf.eset.nouiso.predict.train <- subset(rf.eset.nouiso.predict.train, cancertype == "MCC")

predictions.combined <- rbind(rf.eset.nouiso.predict, rf.eset.nouiso.predict.train)[, -ncol(rf.eset.nouiso.predict.train)]

print(xtable(predictions.combined, digits=3, align='rccccccccc'))
write.csv(predictions.combined, file="tables/tables4.csv")

# avgvotes <- ddply(predictions.combined, .variables="cancertype", colwise(mean, is.numeric))
# 
# rownames(avgvotes) <- c("MCC*", "MCC Variant")
# avgvotes <- avgvotes[, -1]
# 
# print(xtable(avgvotes, digits=3, align='rccccccccc'))
```


<div style="page-break-after:always"></div>

### Table S5: WaGa and Mkl-1 cell line samples (labeled MCC) held out from classifier training (out-of-bag) are classifed as MCC using random forest trained with cell line microarray expression data.

```{r TableS5, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
oobvotes <- data.frame(rf.eset.nouiso$votes,
                       cancertype=rf.eset.nouiso$y)

# oobvotes <- subset(oobvotes, cancertype == "MCC")

avgoobvotes <- ddply(oobvotes, .variables="cancertype", colwise(mean, is.numeric))

avgoobvotes.melted <- melt(avgoobvotes)

avgoobvotes.melted$cancertype <- gsub(" ", ".", avgoobvotes.melted$cancertype)

avgoobvotes.melted$value <- signif(avgoobvotes.melted$value, 2)
avgoobvotes.melted$value <- with(avgoobvotes.melted, ifelse(as.character(cancertype) == as.character(variable), 
                                                            gsub("(.*)",
                                                                 '<div style="font-weight:bold">\\1</div>',
                                                                 value),
                                                            value))

avgoobvotes.formatted <- cast(avgoobvotes.melted, formula=cancertype ~ variable, value="value")
avgoobvotes.formatted <- avgoobvotes.formatted[, -1]
colnames(avgoobvotes.formatted) <- gsub("\\.", " ", colnames(avgoobvotes.formatted))

rownames(avgoobvotes.formatted) <- colnames(avgoobvotes.formatted)


print(xtable(avgoobvotes.formatted, align='rccccccccc'), sanitize.text.function=I)
write.csv(avgoobvotes.formatted, file="tables/tables5.csv")
```

<div style="page-break-after:always"></div>

### Table S6: Short tandem repeat (STR) fingerprinting results for MCC cell lines.

The WaGa, Mkl-1, and UISO cell lines were analyzed with the AmpFlSTR Identifiler PCR Amplification Kit (Applied Biosystems) while the MCC13 and MCC26 cell lines were analyzed with the PowerPlex 18D Kit (Promega).
The two kits analyze the same 16 markers, while the Promega kit includes two additional markers (Penta D and E).

```{r TableS6, results='asis', echo=FALSE}

str.data <- read.table("../data/str.tsv", sep="\t", header=TRUE, stringsAsFactors=FALSE)

rownames(str.data) <- str.data$Marker
str.data <- str.data[, -1]
colnames(str.data) <- c("UISO (early)", "UISO (late)", "WaGa", "Mkl-1", "MCC13", "MCC26")

str.data[str.data == ""] <- "-"

str.xtbl <- xtable(str.data, align=c('r', 'c', 'c', 'c', 'c', 'c', 'c'))

print(str.xtbl, sanitize.text.function=I, include.rownames=TRUE)

write.csv(str.data, file="tables/tables6.csv")
```

<div style="page-break-after:always"></div>

### Table S7: Antibodies used in immunohistochemical analysis.

```{r TableS7, results='asis', echo=FALSE}

antibody.data <- read.csv("../data/antibodies.csv")
colnames(antibody.data) <- c("Antibody", "Company", "Product Number", "Dilution", "Antigen retrieval")
antibody.xtbl <- xtable(antibody.data)

print(antibody.xtbl, include.rownames=FALSE, sanitize.text.function=I)
write.csv(antibody.data, file="tables/tables7.csv")
```

<div style="page-break-after:always"></div>

### Figure S?: qPCR detection of Merkel cell polyomavirus in MCC tumor samples

qPCR was applied to MCC tumor samples in two batches: from 2011-03-16, which included samples `MT01-MT16`, and 2011-03-17, which included samples `MT17-MT24`.
`Actin` was used as a reference gene and the `Mkl-1` cell line was used as a control in both runs.
Plot of the $\Delta \Delta C_t$ for each sample.
Red points indicate samples in which a precise MCV integration has been identified.
Based on these results, the truly negative samples in this cohort are 11, 14, and 16.
All samples were run in duplicate.

```{r setupddct, echo=FALSE, warning=FALSE, message=FALSE}
library(ddCt)

## Set the genes and samples for ddCt computation
gene.levels <- c("MCPVt", "Actin")
name.reference.sample <- c("Mkl-1")
name.reference.gene <- c("Actin")

```

```{r readdataddct, results='asis'}

metadata <- read.csv("..//data/qPCR/metadata.csv")

ctdata.filenames <- list('20110316'="../data/qPCR/qBiomarker_SDS_2011-03-16_14-58-36.txt",
                         '20110317'="../data/qPCR/qBiomarker_SDS_2011-03-17_17-24.txt")

data <- llply(ctdata.filenames, processct, 
              calibrationSample=name.reference.sample, 
              housekeepingGenes=name.reference.gene,
              type="mean")

data.melted <- ldply(data, formatctdata,
                     gene.levels=gene.levels)

colnames(data.melted)[1] <- "Date"

# Remove reference gene
data.melted <- subset(data.melted, !(Gene %in% name.reference.gene))
data.melted$Gene <- factor(data.melted$Gene)

# Merge with clinical data to identify validated MCV samples
data.melted <- merge(data.melted, metadata, by.x="Sample", by.y="sample")


```

```{r plotddct, fig.width=10, fig.height=4}
dodge <- position_dodge(width=0.9)

p <- ggplot(data.melted, subset=!(Gene %in% name.reference.gene),
            aes(color=validated.mcv, x=Sample, y=ddCt))
p <- p + geom_hline(aes(yintercept=1), color="grey")

p <- p + geom_point(size=6)
p <- p + scale_color_manual(values=c("black", "red"))
p <- p + facet_wrap( ~ Date, nrow=1, scales="free_x")

p <- p + labs(y=expression(Delta~Delta~C[t]))
p <- p + theme_bw() + theme(axis.text.x=element_text(size=18, angle=270, vjust=0.5),
                            axis.text.y=element_text(size=18),
                            axis.title=element_text(size=20),
                            legend.position="none")
print(p)
```

